### 2015-05-12 whoop.tex
update in optimisations

A common practice for statically analyzing software is to ``close'' the environment by abstracting away the low level implementation details. Any abstract environmental model, though, can and will ultimately result in decreased precision and thus false positives. This is one of the main limitations of statically analyzing real life software that use complex external APIs. In our case, we developed a simple environmental model for the Linux kernel that consists of stub header files. Because we only focus on lockset-based properties during the race-checking phase of \whoop, we can get away with abstracting a lot of the underlying functionality. Our model, though, can still result in a lot of false positives.

To increase precision we enhance \whoop with domain-specific knowledge of how the Linux kernel uses its various drivers. This is an ongoing manual effort: the more drivers we study, the more domain-specific properties we discover, which we then use for increased precision.

In our experience, one of the most effective precision optimizations is to enrich \whoop with information regarding kernel imposed locks. The Linux kernel is able to serialize a number of entry point calls, which thus can never run concurrently with each other. An an example, a large number of networking and ethernet entry points are mutually serialized with RTNL, a network-specific kernel lock. \whoop will not create pairs for entry points that cannot run concurrently.

Another optimization is to soundly reduce the checked memory locations. If a memory location is accessed only by one entry point in the pair, then it is safe to not check it. This can potentially decrease the amount of generated candidate invariants and speed up the verification process.

Modeling kernel methods is another opportunity to increase precision, but requires expert knowledge about Linux. For example, the \texttt{register\_netdev()} function registers a network driver with the kernel. Before this function is called, the kernel is not able to invoke any of the driver's entry points. Modeling this functionality can reduce false positives when race-checking accesses before the driver's registration.

@@

Finally, we assume that the formal parameters of an entry point do not alias, and thus do not race. This is a potentially unsound feature that can be turned off using a command line option. In our experience so far, though, we have not missed any data races because of this assumption. We also assume that our environmental model and domain-specific knowledge are correct and do not cause any bugs to be missed.

@@

Finally, another limitation of our approach is the need for precise environmental modeling. Enriching \whoop with domain-specific knowledge to increase precision is an ongoing manual effort, but requires expertise in the Linux kernel. We argue that further increasing the precision of our models is orthogonal to the contributions of this paper. Moreover, even if the symbolic lockset analysis can be imprecise, \whoop can still use any race-freedom guarantees to significantly speedup a precise bug-finders, as seen in Sections~\ref{whoop:bugfinding} and~\ref{evaluation}.

>>>>

To increase precision we enhance \whoop with domain-specific knowledge of how the Linux kernel interacts with its drivers. This is an ongoing manual effort: the more drivers we study, the more domain-specific properties we discover, which we then exploit to make \whoop more precise.

In our experience, one of the most effective precision optimizations is to enrich \whoop with information regarding kernel-imposed serialization. The Linux kernel can serialize calls to entry points, thus forcing them to not run concurrently with each other. An an example, a large number of networking and ethernet entry points are mutually serialized with RTNL, a network-specific kernel lock. \whoop exploits this knowledge and does not create pairs for entry points that are serialized.

Modeling kernel methods is another opportunity to increase precision, but requires Linux kernel expertise. For example, the \texttt{register\_netdev()} function registers a network driver with the kernel. Before this function is called, the kernel is not able to invoke any of the driver's entry points. Modeling this functionality can reduce false positives when race-checking accesses before the driver's registration.

Another optimization \whoop uses is to soundly reduce the checked memory locations. If a memory location is accessed only by one entry point in a pair, then it is safe to not check this location for this pair. This reduces the burden to the SMT solver and speed up the verification process.

@@

Finally, we assume that the formal parameters of an entry point do not alias, and thus cannot race. This is a potentially unsound feature that can be turned off using a command line option. In our experience so far, though, we have not missed any data races because of this assumption. We also assume that our environmental model and domain-specific knowledge are correct and do not cause any bugs to be missed.

@@

Statically analyzing system-level software, such as drivers, requires to ``close'' the environment by abstracting away the low level implementation details. In this work, we developed a simple environmental model for the Linux kernel that consists of (nondeterministic) stub functions. A limitation of our model is that it can, and will, ultimately result in decreased precision and, thus, false positives. However, because we currently only focus on finding data races, we can get away with over-approximating a lot of the underlying Linux kernel functionality, without losing too much precision. Enriching our model with domain-specific knowledge to make it more precise is an ongoing manual effort, but requires Linux expertise. We argue that further increasing the precision is orthogonal to the contributions of this paper. Moreover, even if the symbolic lockset analysis results in false positives, \whoop can still use any race-freedom guarantees to significantly speedup a more precise bug-finder, as seen in Sections~\ref{whoop:bugfinding} and~\ref{evaluation}.
